\bf Interactive Visualization of Hierarchically Structured Data

### 2017-03-23 treelapse.tex
Fix some citation appearances

This framework suggests other potential application areas, not all of
which have a priori known tree structures. For example, collections of
spatially-indexed time series are frequently encountered in practice --
consider energy consumption, product sales, or high school dropout rates
across regional districts. This type of data has an implicit tree
structure -- at the top level are different states, while at the bottom
are individual census tracts, say. Analysis here revolves around the
question of how variation across time series is related to their
geographic position. A case study to this type of data is given in
Section \ref{zillow-study}.

Alternatively, if this type of hierarchical contextual information is
not directly available, a tree structure can be learned from the data.
This could be achieved by learning a hierarchical clustering on the
original series. Further, if contextual information is available, but it
is not hierarchical, it is possible to setup a supervised problem that
uses context to predict features of the time series. We can construct a tree by
applying a tree-based classifier \citep{breiman1984classification} or extracting
a regression tree from a more complex supervised model
\citep{boz2002extracting,saito2002extracting}. Analysis then focuses on how
different partitions of the contextual, covariate space relate to observed time
series. This approach is described in Section \ref{bikesharing-study}.

@@

Our third display is directed at the bacterial dynamics question. Here,
two panels are arranged one over the other; one displays all time
series, while the other displays all tree nodes, with node sizes
reflecting the time series value at that node. For this reason, we call
the display, Timebox Trees. In the time series panel, we have directly
implemented the timeboxes idea. We then link the panels: when a set of
series is highlighted by the timeboxes, the associated tree nodes are
also highlighted. For example, timeboxes can be used to focus on a set
of series with specific shape -- increased abundance after an ecological
shock, for example -- and identify along what subtrees this pattern is
present. To further focus on specific elements, a pan-zoom scented
widget is provided \citep{willett2007scented}. The widget is a miniature version
of the full time series panel, equipped with a single brush whose extent
specifies the limits in the main time series panel. As in the DOI trees and
sankeys, a search bar can be used to highlight those series of interest a
priori.

The final display currently implemented in the package is the natural
converse of the timebox trees display. Rather than defining visual
queries in terms of time series, it defines queries using nodes in the
tree. For this reason, we call the display Treeboxes. Rather than
focusing on the intersection of brushes, as in timebox trees, we focus
on the union of brushed over nodes. This allows us to highlight series
associated with nodes on distant subtrees. This display is also suited
for the bacterial dynamics problem. For example, by highlighting all
nodes at one taxonomic level in the tree, we can easily summarize the
time series pattern for all the taxa at that level. Alternatively,
focusing on all the children below a single node makes it possible to
see how much correlation and competition there is between taxonomically
similar bacteria. As in the timebox trees display, a search box and
pan-zoom scented widget are provided.

@@

Therefore, our focus here is on the differential abundance question,
rather than dynamics. We would like to provide visual representations of
differential abundance across CSTs and also between preterm and non-preterm
births. \citet{digiulio2015temporal} interpreted the CSTs using a heatmap, with
bacteria ordered according to a hierarchical clustering. By using the DOI sankey
instead, we can interpret the CSTs in their taxonomic context and at multiple
scales of taxonomic resolution. Further, while 
\citep{digiulio2015temporal} focused on identifying associations between
preterm births and CSTs -- presumably because testing individual bacteria loses
power -- we can compare bacterial abundances between preterm and non-preterm
samples along subtrees, without requiring CSTs as an intermediary.

@@

We next consider an application unrelated to the microbiome, but with
relatively clear hierarchical structure. Our data are downloaded from
Zillow\footnote{\url{http://www.zillow.com/research/data/}}, and give
the Zillow Home Value Indexes at the neighborhood level, across the
country, computed monthly between 1996 and 2016. In our display, we have
taken the base 10 log of these indexes. As our hierarchical structure,
we use each neighborhood's assignment to state, regional, county, and
city levels. We represent each of these coarser spatial categories using
the average of all neighborhoods contained in them. We have filtered
down to the 890 neighborhoods in California; rendering more
neighborhoods while keeping all 246 timepoints causes the interface to
lag\footnote{See Section \ref{conclusion} for potential optimizations,
  however.}.

@@

\bibliographystyle{plain}
\bibliography{bibliography}

>>>>

This framework suggests other potential application areas, not all of
which have a priori known tree structures. For example, collections of
spatially-indexed time series are frequently encountered in practice --
consider energy consumption, product sales, or high school dropout rates
across regional districts. This type of data has an implicit tree
structure -- at the top level are different states, while at the bottom
are individual census tracts, say. Analysis here revolves around the
question of how variation across time series is related to their
geographic position.

Alternatively, if this type of hierarchical contextual information is
not directly available, a tree structure can be learned from the data.
This could be achieved by learning a hierarchical clustering on the
original series. Further, if contextual information is available, but it
is not hierarchical, it is possible to setup a supervised problem that
uses context to predict features of the time series. We can construct a tree by
applying a tree-based classifier \citep{breiman1984classification} or extracting
a regression tree from a more complex supervised model
\citep{boz2002extracting,saito2002extracting}. Analysis then focuses on how
different partitions of the contextual, covariate space relate to observed time
series.

@@

Our third display is directed at the bacterial dynamics question. Here,
two panels are arranged one over the other; one displays all time
series, while the other displays all tree nodes, with node sizes
reflecting the time series value at that node. For this reason, we call
the display, timebox trees. In the time series panel, we have directly
implemented the timeboxes idea. We then link the panels: when a set of
series is highlighted by the timeboxes, the associated tree nodes are
also highlighted. For example, timeboxes can be used to focus on a set
of series with specific shape -- increased abundance after an ecological
shock, for example -- and identify along what subtrees this pattern is
present. To further focus on specific elements, a pan-zoom scented
widget is provided \citep{willett2007scented}. The widget is a miniature version
of the full time series panel, equipped with a single brush whose extent
specifies the limits in the main time series panel. As in the DOI trees and
sankeys, a search bar can be used to highlight those series of interest a
priori.

The final display currently implemented in the package is the natural
converse of the timebox trees display. Rather than defining visual
queries in terms of time series, it defines queries using nodes in the
tree. For this reason, we call the display treeboxes. Rather than
focusing on the intersection of brushes, as in timebox trees, we focus
on the union of brushed over nodes. This allows us to highlight series
associated with nodes on distant subtrees. This display is also suited
for the bacterial dynamics problem. For example, by highlighting all
nodes at one taxonomic level in the tree, we can easily summarize the
time series pattern for all the taxa at that level. Alternatively,
focusing on all the children below a single node makes it possible to
see how much correlation and competition there is between taxonomically
similar bacteria. As in the timebox trees display, a search box and
pan-zoom scented widget are provided.

@@

Therefore, our focus here is on the differential abundance question,
rather than dynamics. We would like to provide visual representations of
differential abundance across CSTs and also between preterm and non-preterm
births. \citet{digiulio2015temporal} interpreted the CSTs using a heatmap, with
bacteria ordered according to a hierarchical clustering. By using the DOI sankey
instead, we can interpret the CSTs in their taxonomic context and at multiple
scales of taxonomic resolution. Further, while
\citet{digiulio2015temporal} focused on identifying associations between
preterm births and CSTs -- presumably because testing individual bacteria loses
power -- we can compare bacterial abundances between preterm and non-preterm
samples along subtrees, without requiring CSTs as an intermediary.

@@

We next consider an application unrelated to the microbiome, but with
relatively clear hierarchical structure. Our data are downloaded from
Zillow\footnote{\url{http://www.zillow.com/research/data/}}, and give
the Zillow Home Value Indexes at the neighborhood level, across the
country, computed monthly between 1996 and 2016. In our display, we have
taken the base 10 log of these indexes. As our hierarchical structure,
we use each neighborhood's assignment to state, regional, county, and
city levels. We represent each of these coarser spatial categories using
the average of all neighborhoods contained in them. We have filtered
down to the 890 neighborhoods in California; rendering more
neighborhoods while keeping all 246 timepoints causes the interface to
lag\footnote{See below for potential optimizations, however.}.

@@

\bibliographystyle{plainnat}
\bibliography{bibliography}
